---
title: "causalForest - log maize yields placebo test"
author: "Jill Deines"
date: "8/6/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**Goal: run causal forest with mean 2001-2010 yields, as a placebo test of the method**

Notes/Updates

* log scale
* reduced variables

**note for published repository:**
data provided in this repository have had latitude/longitude removed to preserve data privacy; since latitude and longitude were used as variables in the published analysis, this code will not reproduce the results exactly.

publication output is included as data files for use in scripts to recreate the results figures with the official analysis output. see scripts 3.05-03.35.


This also pulls in output from the main maize causal forests, which is not supplied here since that result depends on/includes lat lon.

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, fig.path='../../figure/02.07_causalForest_maize_log2_placebo/')
```

**R Packages Needed**

```{r packages, message= F, echo = T, eval = T}
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(grf)
library(RColorBrewer)
library(earth)
library(corrplot)
library(car)
library(aod)
library(sf)
library(Rcpp)

library(here)


sessionInfo()
```

**Directories**

```{r setDirs}
# input/cleaned data folder
projDir <- here::here()
dataDir <- paste0(projDir,'/data/cleanedPointSample')
dataFile <- 'pointSample_kaiyuV2_20220127_maize_placeboYields_latLonRemoved.csv'

# scratch folder for model rdata objects
scratchFolderMain <- '/Users/deinesji/Dropbox/2Stanford/projects/2021_coverCrops/data/causalForests/20220201_scratch_6state_maize_v04_log_lessVars'
scratchFolder <- '/Users/deinesji/Dropbox/2Stanford/projects/2021_coverCrops/data/causalForests/20220201_scratch_6state_maize_v04_log_lessVars_placebo'

# make scratch folder if necessary
dir.create(file.path(scratchFolder), showWarnings = FALSE)
```

# Load and clean
load master dataset created in masterDatasetCausalFOrest.rmd, with yields replaced with the 2001-2010 mean

```{r load}

CoverCropYearsAll_placebo <- read_csv(paste0(dataDir, '/', dataFile))
```


# Causal Forest


## causal forest

### filter dataset by overlap in original test

```{r getOverlap}
overlap <- readRDS(paste0(scratchFolderMain, '/overlap.rds'))

placeboOverlap <- CoverCropYearsAll_placebo %>%
  left_join(overlap %>% dplyr::select(uniqueID, w_hat)) %>%
  filter(uniqueID %in% overlap$uniqueID) %>%
  filter(!is.na(yield_tha))

# y_hat and reduce to correct dataset
Y_hat_main <- readRDS(paste0(scratchFolderMain, '/Y_hat.rds'))
overlap$Y_hat <- Y_hat_main
Y_hat <- overlap %>%
  filter(uniqueID %in% placeboOverlap$uniqueID) %>%
  pull(Y_hat)
```

```{r causal, eval = FALSE}

# reload mars variables
marsAll <- readRDS(paste0(scratchFolderMain, '/mars_c1.rds'))
ev <- evimp(marsAll) # estimate variable importance
envvars <- rownames(ev)

# treatment
W <- placeboOverlap %>% dplyr::pull(W)
w_hat <- placeboOverlap$w_hat

# outcome
Y <- placeboOverlap %>% dplyr::pull(Y)

# make covariate list; 
# add in important propensity variables
varsToKeep_main = c(envvars, 'year', 'slope')

Xmain <- placeboOverlap %>%
  dplyr::select(all_of(varsToKeep_main) )

cf <- grf::causal_forest(Xmain, Y, W, Y_hat, w_hat, num.trees = 2000)

varImpMain <- cf %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(cf$X.orig)) %>% 
  arrange(desc(V1))
varImpMain

saveRDS(cf, paste0(scratchFolder, '/cf_clusters_xc1.rds'))
saveRDS(varImpMain, paste0(scratchFolder, '/cf_clusters_xc1_varImp.rds'))
```


# Checking things out

reload and extract model outputs

```{r reload}
# data used - placeboOverlap


# propensity model results
W_hat <- placeboOverlap %>% dplyr::pull(w_hat)

# causal forest
cf <- readRDS(paste0(scratchFolder, '/cf_clusters_xc1.rds'))
varImp <- readRDS(paste0(scratchFolder, '/cf_clusters_xc1_varImp.rds'))


W <- placeboOverlap %>% dplyr::pull(W)
Y <- placeboOverlap %>% dplyr::pull(Y)


# process results ------------------------------
# add predicted treatment effects to data 
constOob_predict <- predict(cf)

# add hats to original data frame
placeboOverlapW0 <- placeboOverlap %>%
  bind_cols(data.frame(Y_hat = Y_hat))


colVars <- varImp %>% dplyr::pull(variable)

placeboOverlapW <- placeboOverlapW0 %>%
  mutate_at(colVars, list(Q4 = ~ntile(., 4),
                          Q5 = ~ntile(., 5)))

overlapTau <- placeboOverlapW %>%
  bind_cols(constOob_predict) 
```

## Average treatment Effects

```{r ate}
varImp

ate_cf_aipw = average_treatment_effect(cf)
ate_cf_aipw = average_treatment_effect(cf)
tauhat_rf_aipw = c(ATE=ate_cf_aipw["estimate"],
                   lower_ci=ate_cf_aipw["estimate"] - 1.96 * ate_cf_aipw["std.err"],
                   upper_ci=ate_cf_aipw["estimate"] + 1.96 * ate_cf_aipw["std.err"])

tauhat_rf_aipw

ggplot(overlapTau,
       aes(x = predictions)) +
  geom_histogram() +
  #xlim(-.2,.05)+
  xlab('CATE (log t/ha)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = ate_cf_aipw["estimate"], col = 'red') +
  theme_bw()

```

cate by percent

```{r cate_percent}

tauhat_rf_aipw_percent <- (exp(tauhat_rf_aipw) -1) * 100
tauhat_rf_aipw_percent

cate_percent <-  overlapTau %>%
  mutate(tauPercents =  (exp(predictions) -1) * 100,) 

ggplot(cate_percent,
       aes(x = tauPercents)) +
  geom_histogram() +
  #xlim(-13,2.5)+
  xlab('CATE (% of mean log yield)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent['ATE.estimate'], col = 'red') +
  theme_bw()

ggplot(cate_percent,
       aes(x = tauPercents)) +
  geom_histogram() +
  #xlim(-13,2.5)+
  xlab('CATE (% of mean yield)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent['ATE.estimate'], col = 'red') +
  theme_bw()+
  facet_wrap(~state)

ggplot(cate_percent,
       aes(x = tauPercents)) +
  geom_histogram() +
 # xlim(-13,2.5)+
  xlab('CATE (% of mean yield)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent['ATE.estimate'], col = 'red') +
  theme_bw()+
  facet_wrap(~year)
```

