---
title: "causalForest - soy log"
author: "Jill Deines"
date: "2/1/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**Goal: run and evaluate causal forest on soy fields in IN, IL, IA, OH, MO, MI for 2019 & 2020**

Notes/Updates

* log scale
* reduced variables


**note for published repository:**
data provided in this repository have had latitude/longitude removed to preserve data privacy; since latitude and longitude were used as variables in the published analysis, this code will not reproduce the results exactly.

publication output is included as data files for use in scripts to recreate the results figures with the official analysis output. see scripts 3.05-03.35.

```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, fig.path='../../figure/02.22_causalForest_soy_log2/')
```

**R Packages Needed**

```{r packages, message= F, echo = T, eval = T}
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(grf)
library(RColorBrewer)
library(earth)
library(corrplot)
library(car)
library(aod)
library(sf)
library(Rcpp)

library(here)


sessionInfo()
```

**Directories**

```{r setDirs}
# input/cleaned data folder
projDir <- here::here()
dataDir <- paste0(projDir,'/data/cleanedPointSample')
dataFile <- 'pointSample_kaiyuV2_20220127_soybeans_latLonRemoved.csv'

# scratch folder for model rdata objects
scratchFolder <- '/Users/deinesji/Dropbox/2Stanford/projects/2021_coverCrops/data/causalForests/20220127_scratch_6state_SOY_v04_log_lessVars'

# make scratch folder if necessary
dir.create(file.path(scratchFolder), showWarnings = FALSE)
```

# Load and clean
load master dataset created in masterDatasetCausalFOrest.rmd

```{r loadData}
masterDataset <- read_csv(paste0(dataDir, '/', dataFile), guess_max = 10000)

# remove extreme yield values
cleanQ <- quantile(masterDataset$yield_tha, probs = c(0.0001, 0.9999), na.rm=TRUE)
cleanQ          

CoverCropYearsAll <- masterDataset %>%
  ungroup() %>%
  filter(yield_tha > cleanQ[1]) %>%
  filter(yield_tha < cleanQ[2]) %>%
  mutate(uniqueID = row_number()) %>%
  mutate(Y = log(yield_tha))

hist(CoverCropYearsAll$Y)
hist(CoverCropYearsAll$yield_tha)


table(CoverCropYearsAll$W)
```



# Causal Forest

## Propoensity model
retain lat lon variables

```{r propensity, eval = FALSE}

X_norm <- CoverCropYearsAll %>%
  dplyr::select(c(contains('norm'), "nccpicorn",'drclasscd','rootznaws', 'slope',
                  'latitude','longitude'))

W <- CoverCropYearsAll %>% dplyr::pull(W)

W_forest <- grf::regression_forest(X_norm, W, num.trees = 2000)

W_hat <- predict(W_forest)$predictions

WvarImp <- W_forest %>% 
  grf::variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(W_forest$X.orig)) %>% 
  arrange(desc(V1))

# save intermediates
saveRDS(CoverCropYearsAll, paste0(scratchFolder, '/CoverCropYearsAll.rds'))
saveRDS(W_forest, paste0(scratchFolder, '/W_forest.rds'))
saveRDS(W_hat, paste0(scratchFolder, '/W_hat.rds'))
saveRDS(WvarImp, paste0(scratchFolder, '/W_forest_varImp.rds'))
```

### Prune samples
Toss out rows with excessively high/low propensities

```{r propensityFilter}
# re-load so chunk runs in markdown output
#CoverCropYearsAll <- readRDS(paste0(scratchFolder, '/CoverCropYearsAll.rds'))

W <- CoverCropYearsAll %>% dplyr::pull(W)

W_hat <- readRDS(paste0(scratchFolder, '/W_hat.rds'))
WvarImp <- readRDS(paste0(scratchFolder, '/W_forest_varImp.rds'))
WvarImp

# check propensity scores
p_rf = W_hat
hist(p_rf, xlab = 'Propensity Scores', main = 'Histogram of Propensity Scores Corn')

min(p_rf)
sum(p_rf < .05)

{plot(smooth.spline(p_rf, W, df = 4))
abline(0, 1)}

# filter it
overlap0 <- CoverCropYearsAll %>%
  mutate(w_hat = W_hat) %>%
  mutate(keep = case_when(w_hat < 0.05 | w_hat > 0.95 ~ 0,
                        w_hat >= 0.05 | w_hat <= 0.95  ~ 1))
overlap <- overlap0 %>% filter(keep == 1)
nrow(overlap0) - nrow(overlap)
nrow(overlap)

both <- data.frame(W_hat = W_hat,
                   type = paste0('original sampling (',nrow(CoverCropYearsAll), ' obs)')) %>%
  bind_rows(data.frame(W_hat = overlap$w_hat,
                       type = paste0('Propensity Filtered (', nrow(overlap), ' obs)')))

ggplot(both,
       aes(x = W_hat)) +
  geom_histogram() +
  geom_vline(xintercept = 0.05, col='red') +
  geom_vline(xintercept = 0.95, col = 'red') +
  facet_wrap(~type, nrow = 2) +
  #ggeasy::easy_center_title() +
  ggtitle('Soy') +
  theme_bw()

table(overlap$W)
table(overlap$year)
table(overlap[,c('W','year')])

table(overlap[,c('state')])
table(overlap[,c('state',"W")])

# save overlap data!
saveRDS(overlap, paste0(scratchFolder, '/overlap.rds'))
```


## Expected Outcome: y.hat

### Pick covars
And make sure to take out yield_tha!! ha


Here, removes gldas variables (redundant with tc soilm) and tc_def (redundant with vpd)

```{r mars, eval = FALSE}
overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))

# subset of variables to choose from
X_c1 <- overlap %>%
  dplyr::select(-c (W,  uniqueID, fips5, w_hat, keep, cropType,contains('norm'),
                    latitude, longitude, mukey, state_fips, fips, phinge, vhinge,
                    state_name, state, ccFreq, nccpicorn, yield_tha,
                    contains('gldas'), contains('tc_def'))) %>%
  tidyr::drop_na()

# mars for var dplyr::selection
set.seed(5)
marsAll <- earth::earth(Y ~ ., data = X_c1, degree = 2)

saveRDS(marsAll, paste0(scratchFolder, '/mars_c1.rds'))
```

### Train Yhat

```{r Yhat, eval = FALSE}
# variables from mars
marsAll <- readRDS(paste0(scratchFolder, '/mars_c1.rds'))
summary(marsAll)
ev <- evimp(marsAll) # estimate variable importance
ev

envvars <- rownames(ev)
varsToKeep = c(envvars, 'year')

X <- overlap %>%
  dplyr::select(all_of(varsToKeep)) 

corr <- cor(X )
corrplot(corr, type = 'upper')

Y <- overlap %>% dplyr::pull(Y)

Y_forest <- grf::regression_forest(X, Y, num.trees = 2000)

YvarImp <- Y_forest %>% 
  grf::variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(Y_forest$X.orig)) %>% 
  arrange(desc(V1))
YvarImp

Y_hat <- predict(Y_forest)$predictions



saveRDS(Y_hat, paste0(scratchFolder, '/Y_hat.rds'))
saveRDS(Y_forest, paste0(scratchFolder, '/Y_forest.rds'))
saveRDS(YvarImp, paste0(scratchFolder, '/Y_forest_varImp.rds'))
```

### evaluate outcome model

```{r outcomeEval}
overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))
Y_hat <- readRDS(paste0(scratchFolder, '/Y_hat.rds'))
YvarImp <- readRDS(paste0(scratchFolder, '/Y_forest_varImp.rds'))

YvarImp
Y <- overlap %>% dplyr::pull(Y)

### Outcome Estimation Y_Hat
Ychecker <- data.frame(Y = Y,
                       Y_hat = Y_hat,
                       year = overlap$year)
Yagreement <- lm(Y~Y_hat)
summary(Yagreement)

# yield year by year
YcheckerLong <- Ychecker %>%
  tidyr::gather(., key = variable, value = value, Y:Y_hat)


ggplot(YcheckerLong, aes(x=as.factor(year), y = value, group = interaction(year,variable), 
                         fill = variable)) +
  geom_boxplot() +
  theme_bw() +
  ylab('Yield (t/ha)') +
  xlab('Year')

ggplot(Ychecker, aes(x=Y, y = Y_hat)) +
  geom_hex() +
  scale_fill_gradientn(colors = brewer.pal(9,'YlGnBu')) +
  geom_smooth(method= 'lm') +
  coord_equal(xlim=c(0,2), ylim=c(0,2)) + 
  geom_abline(slope = 1, intercept = 0, linetype='dashed') + 
  facet_wrap(~year) +
  theme_bw()
```

## causal forest

```{r causal, eval = FALSE}
# reload overlap
overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))
Y_hat <- readRDS(paste0(scratchFolder, '/Y_hat.rds'))

# reload mars variables
marsAll <- readRDS(paste0(scratchFolder, '/mars_c1.rds'))
ev <- evimp(marsAll) # estimate variable importance
envvars <- rownames(ev)

# treatment
W <- overlap %>% dplyr::pull(W)
w_hat <- overlap$w_hat

# outcome
Y <- overlap %>% dplyr::pull(Y)

# make covariate list; 
# add in important propensity variables
varsToKeep_main = c(envvars,  'year','slope','drclasscd')

Xmain <- overlap %>%
  dplyr::select(all_of(varsToKeep_main) )
  

corr <- cor(Xmain )
corrplot(corr, type = 'upper')


cf <- grf::causal_forest(Xmain, Y, W, Y_hat, w_hat, num.trees = 2000)


varImpMain <- cf %>% 
  variable_importance() %>% 
  as.data.frame() %>% 
  mutate(variable = colnames(cf$X.orig)) %>% 
  arrange(desc(V1))


saveRDS(cf, paste0(scratchFolder, '/cf_clusters_xc1.rds'))
saveRDS(varImpMain, paste0(scratchFolder, '/cf_clusters_xc1_varImp.rds'))
```


# Checking things out

reload and extract model outputs

```{r reload}
# data used
overlap <- readRDS(paste0(scratchFolder, '/overlap.rds'))

# propensity model results
W_hat <- overlap %>% dplyr::pull(w_hat)
WvarImp <- readRDS(paste0(scratchFolder, '/W_forest_varImp.rds'))
W_hat_all <- readRDS(paste0(scratchFolder, '/W_hat.rds'))

# outcome model
Y_hat <- readRDS(paste0(scratchFolder, '/Y_hat.rds'))
YvarImp <- readRDS(paste0(scratchFolder, '/Y_forest_varImp.rds'))

# causal forest
cf <- readRDS(paste0(scratchFolder, '/cf_clusters_xc1.rds'))
varImp <- readRDS(paste0(scratchFolder, '/cf_clusters_xc1_varImp.rds'))


# reconstruct subsets
X_norm <- CoverCropYearsAll %>%
  dplyr::select(c(contains('norm'), "nccpisoy",'drclasscd','rootznaws',  'latitude','longitude','slope'))

W <- overlap %>% dplyr::pull(W)
Y <- overlap %>% dplyr::pull(Y)


# process results ------------------------------
# add predicted treatment effects to data 
constOob_predict <- predict(cf)

# add hats to original data frame
overlapW0 <- overlap %>%
  bind_cols(data.frame(Y_hat = Y_hat))

colNormals <- names(X_norm)
colVars <- varImp %>% dplyr::pull(variable)

overlapW <- overlapW0 %>%
  mutate_at(colNormals, list(Q4n = ~ntile(., 4),
                             Q5n = ~ntile(., 5))) %>%
  mutate_at(colVars, list(Q4 = ~ntile(., 4),
                          Q5 = ~ntile(., 5)))

overlapTau <- overlapW %>%
  bind_cols(constOob_predict) 
```

## Propensity by covariates

```{r propensity_trends, fig.width = 7, fig.height = 6.5, dpi = 300}
# add hats to original data frame
prop_all <- CoverCropYearsAll %>%
  bind_cols(data.frame(W_hat = W_hat_all))

# calculate quantiles of interest
colNormals <- names(X_norm)
colVars <- varImp %>% dplyr::pull(variable)

prop_all2 <- prop_all %>%
  mutate_at(colNormals, list(Q4n = ~ntile(., 4),
                           Q5n = ~ntile(., 5))) %>%
  mutate_at(colVars, list(Q4 = ~ntile(., 4),
                           Q5 = ~ntile(., 5)))

# plot normals vs What : quantiles
propensityVarsLong2 <- prop_all2 %>%
  dplyr::select(c(W_hat, contains('Q4n'))) %>%
  tidyr::gather(., key = variable, value = value, pr_apr_norm_mm_Q4n:slope_Q4n)

ggplot(propensityVarsLong2,
       aes(y = W_hat, x = value, group = value)) +
  geom_boxplot() +
  facet_wrap(~variable) +
  theme_bw() + xlab('Quartile') + ylab('W.hat = conditional prob of treatment')
```



## Average treatment Effects

```{r ate}
varImp

ate_cf_aipw = average_treatment_effect(cf)
ate_cf_aipw = average_treatment_effect(cf)
tauhat_rf_aipw = c(ATE=ate_cf_aipw["estimate"],
                   lower_ci=ate_cf_aipw["estimate"] - 1.96 * ate_cf_aipw["std.err"],
                   upper_ci=ate_cf_aipw["estimate"] + 1.96 * ate_cf_aipw["std.err"])

tauhat_rf_aipw

ggplot(overlapTau,
       aes(x = predictions)) +
  geom_histogram() +
 xlim(-.2,.05)+
  xlab('CATE (log t/ha)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = ate_cf_aipw["estimate"], col = 'red') +
  theme_bw()


```

### percentage

```{r cate_percent}
tauhat_rf_aipw_percent <- (exp(tauhat_rf_aipw) -1) * 100
tauhat_rf_aipw_percent

cate_percent <-  overlapTau %>%
  mutate(tauPercents = (exp(predictions) -1) * 100) 

ggplot(cate_percent,
       aes(x = tauPercents)) +
  geom_histogram() +
  xlim(-13,2.5)+
  xlab('CATE (% of mean log yield)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent['ATE.estimate'], col = 'red') +
  theme_bw()

ggplot(cate_percent,
       aes(x = tauPercents)) +
  geom_histogram() +
  xlim(-13,2.5)+
  xlab('CATE (% of mean yield)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent['ATE.estimate'], col = 'red') +
  theme_bw()+
  facet_wrap(~state)

ggplot(cate_percent,
       aes(x = tauPercents)) +
  geom_histogram() +
  xlim(-13,2.5)+
  xlab('CATE (% of mean yield)') +
  geom_vline(xintercept = 0, col = 'black', linetype = 'dashed') +
  geom_vline(xintercept = tauhat_rf_aipw_percent['ATE.estimate'], col = 'red') +
  theme_bw()+
  facet_wrap(~year)
```

##  HTE

```{r hte_test}
# compare regions with high and low CATES 
high_effect = overlapTau$predictions > median(overlapTau$predictions)
ate.high = average_treatment_effect(cf, subset = high_effect, target.sample = 'overlap')
ate.low = average_treatment_effect(cf, subset = !high_effect, target.sample = 'overlap')

paste("95% CI for difference in ATE:", round(ate.high[1] - ate.low[1], 3), "+/-", round(qnorm(0.975) * sqrt(ate.high[2]^2 + ate.low[2]^2), 3))

# run best linear predictor analysis
calibration <- test_calibration(cf)
calibration
```

## effects by covariates

```{r hte_vars_all_catebin}
hte_vars <- colVars[!(colVars %in% c('year'))]

overlapTau_1 <- overlapTau %>%
   mutate(cate_ntile = ntile(predictions, 3))

ggplot(overlapTau_1,
       aes(x = cate_ntile, group = cate_ntile, y = predictions)) +
  geom_boxplot()

hte_df <- overlapTau_1 %>%
  dplyr::select(c(uniqueID, year, predictions, cate_ntile, all_of(hte_vars))) %>%
  tidyr::gather(., key = variable, value = value, hte_vars)

# plot distribution of covariates by cate bins
ggplot(hte_df,
       aes(x = cate_ntile, y = value, group = cate_ntile)) +
  geom_boxplot() + 
  facet_wrap(~variable, scales = 'free_y', nrow = 4) +
  theme_bw() + 
  ggtitle('Covariate values by CATE bins')
```


