---
title: "Cover Crop Trends - Figure"
author: "Jill Deines"
date: "1/28/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

**Goal: Look at cover crop trends over time from the RS dataset, along with NASS Census**


Notes:
- makes Figure 2 and Fig 1c
- input data files exported from GEE in script users/jdeines/Lobell_coverCropII/00.10_formatCcMaps_getAreaByState
- county level NASS data from Qu. counties in CC study from GEE script 00.20 


```{r knitrOpts, echo=FALSE}
library(knitr)
opts_chunk$set(cache=FALSE, fig.path='../../figure/03.02_coverCropTrends/')
```

**R Packages Needed**

```{r packages, message=FALSE, echo=TRUE, eval=TRUE}
library(tidyverse)
library(here)
library(tidycensus)
library(patchwork)

sessionInfo()
```

**Directories**

```{r setDirs}
projDir <- here::here()
dataDir <- paste0(projDir,'/data/stateTrends_RS_nass')

# remote sensing tabulations
outFile2 <- 'stateSummary_kaiyuV2_noAgMask_2000_2020.csv'

# county NASS data - cover crop maps
countyFile <- paste0(dataDir,'/2012_2017_NASS_Cover_Crop_Stats_countyFromQu.csv')
countyFilter <- read_csv(paste0(dataDir,'/StudyAreaCountyList_CCmapPresence_gt5000.csv'))

# county NASS data - ag area planted (compiled in 00.40 getNassData)
nassA0_file <- paste0(dataDir,'/nass_countyAgArea/NASS_cornBelt6_countyAgArea_2020-2021_formatted.csv')

studystates <- c('IA','IN', 'IL', 'OH', 'MO','MI')

data(fips_codes)
stateNames <- fips_codes %>%
  dplyr::group_by(state_code) %>%
  slice(1) %>% ungroup() %>%
  dplyr::select(state, state_name, state_code) %>%
  rename(STATE_ABBR = state)

```


# cc by state - Fig 1c

```{r stateCheck}
studystates2 <- c(studystates, 'WI','MN','SD')
nass0 <- read_csv(countyFile)

nass_check <- nass0 %>%
  # conversions
  mutate(totalAg_ha_2012 = `Cropland in 12 (arce)` / 2.471,
         totalAg_ha_2017 = `Cropland in 17 (arce)` / 2.471,
         coverCropped_ha_2012 = `CoverCrop in 12 (arce)` / 2.471,
         coverCropped_ha_2017 = `CoverCrop in 17 (acre)` / 2.471,
         state_code = substr(as.character(FIPS),1,2))  %>%
  group_by(state_code) %>%
  summarize(totalAg_ha_2012 = sum(totalAg_ha_2012, na.rm = TRUE),
         totalAg_ha_2017 = sum(totalAg_ha_2017, na.rm = TRUE),
         coverCropped_ha_2012 = sum(coverCropped_ha_2012, na.rm = TRUE),
         coverCropped_ha_2017 = sum(coverCropped_ha_2017, na.rm = TRUE)) %>%
  left_join(stateNames) %>%
  filter(STATE_ABBR %in% studystates2) %>%
  tidyr::gather(., key = variable, value = value, contains('20')) %>%
  mutate(Year = as.integer(str_sub(variable, -4, -1)),
         variable = str_sub(variable, 1, -6)) %>%
  tidyr::spread(., key = variable, value = value) %>%
  mutate(percentCoverCrop = coverCropped_ha/totalAg_ha * 100,
         source = 'NASS') 


# total cover crop fraction - study area only!
nass_check %>%
  filter(STATE_ABBR %in% studystates) %>%
  group_by(Year) %>% 
  summarize(coverCropped_ha = sum(coverCropped_ha),
            totalAg_ha = sum(totalAg_ha)) %>%
  mutate(percent = coverCropped_ha/totalAg_ha * 100)
```

## 2017 area plot - ms

```{r Fig1c_2017NassBar, fig.width = 2.85, fig.height = 2.5, dpi = 300, dev = c('png','pdf')}
nass_areaPlot <- nass_check %>% 
  filter(Year == 2017 & STATE_ABBR != 'SD') %>%
  mutate(inStudy = case_when(STATE_ABBR %in% c('WI','MN') ~ 0,
                             !(STATE_ABBR %in% c('WI','MN')) ~ 1))

ggplot(nass_areaPlot,
       aes(y = coverCropped_ha/1000, x = reorder(STATE_ABBR, -coverCropped_ha), 
           group = STATE_ABBR, fill = as.factor(inStudy))) +
  geom_bar(stat = 'identity') +
  scale_fill_manual(values = c('gray70','cornflowerblue')) +
  ylab('Cover Crop Area (1000 ha)') + xlab('State') +
  theme_bw() + theme(legend.position = 'none')
```


# Fig 2

## load GEE and nass

```{r loadData}

# load/format NASS
nass0 <- read_csv(countyFile)
nass <- nass0 %>%
  # filter by counties with CC map data
  filter(FIPS %in% countyFilter$GEOID) %>%
  # conversions
  mutate(totalAg_ha_2012 = `Cropland in 12 (arce)` / 2.471,
         totalAg_ha_2017 = `Cropland in 17 (arce)` / 2.471,
         coverCropped_ha_2012 = `CoverCrop in 12 (arce)` / 2.471,
         coverCropped_ha_2017 = `CoverCrop in 17 (acre)` / 2.471,
         state_code = substr(as.character(FIPS),1,2))  %>%
  group_by(state_code) %>%
  summarize(totalAg_ha_2012 = sum(totalAg_ha_2012, na.rm = TRUE),
         totalAg_ha_2017 = sum(totalAg_ha_2017, na.rm = TRUE),
         coverCropped_ha_2012 = sum(coverCropped_ha_2012, na.rm = TRUE),
         coverCropped_ha_2017 = sum(coverCropped_ha_2017, na.rm = TRUE)) %>%
  left_join(stateNames) %>%
  filter(STATE_ABBR %in% studystates) %>%
  tidyr::gather(., key = variable, value = value, contains('20')) %>%
  mutate(Year = as.integer(str_sub(variable, -4, -1)),
         variable = str_sub(variable, 1, -6)) %>%
  tidyr::spread(., key = variable, value = value) %>%
  mutate(percentCoverCrop = coverCropped_ha/totalAg_ha * 100,
         source = 'NASS')


# recalculate percent with new NASS area

# nass area - for unmasked percents (downloaded for work with Brian)
nassA0 <- read_csv(nassA0_file) %>%
  filter(fips5 %in% countyFilter$GEOID) %>%
  group_by(year, state_alpha) %>%
  summarize(Nass_Area_m2 = sum(Nass_Area_m2, na.rm = TRUE)) %>%
  mutate(totalAg_ha_filter = Nass_Area_m2 / 10000) %>%
  rename(STATE_ABBR = state_alpha,
         Year = year) %>%
  dplyr::select(c(STATE_ABBR, Year, totalAg_ha_filter)) %>%
  ungroup()


stateTotals_nassBase <- read_csv(paste0(dataDir, '/', outFile2)) %>%   
  filter(STATE_ABBR %in% studystates) %>%
  left_join(nassA0) %>%
  mutate(percentCoverCrop = coverCropped_ha / totalAg_ha_filter * 100)

```


# Figures

```{r Fig_manu, fig.width = 7, fig.height = 3.55, dpi = 300, dev = c('png','pdf')}
stateLabels <- nass %>% group_by(state_name) %>% slice(1)

p_area <- ggplot(stateTotals_nassBase  ,
       aes(x = Year, y = coverCropped_ha/1000, group = STATE_ABBR)) +
  geom_line() +
      geom_text(data = stateLabels, x = 2001, y = 400, size = 3.5, color = 'black',
                hjust = 'left',
            aes(label = state_name)) +
    geom_point(data = nass,
               aes(x = Year, y = coverCropped_ha/1000), color = 'red') +
  ylab('Area (1000 ha)') +
      facet_wrap(~state_name, nrow = 1) +
  theme_bw() + theme(strip.background = element_blank(),
                     strip.text = element_blank(),
                     axis.title.x = element_blank(),
                    # axis.text.x = element_blank(),
                    #panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())


p_percent <- ggplot(stateTotals_nassBase,
       aes(x = Year, y = percentCoverCrop, group = STATE_ABBR)) +
  geom_line() +
  geom_text(data = stateLabels, x = 2001, y = 14.5, size = 3.5, color = 'black',
                hjust = 'left',
            aes(label = state_name)) +
  geom_point(data = nass ,
             aes(x = Year, y = percentCoverCrop), color = 'red') +
  ylab('% of Total Area Planted') +
  facet_wrap(~state_name, nrow = 1) +
  theme_bw() + theme(strip.background = element_blank(),
                     strip.text = element_blank(),
                    #axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                    #panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())




p_area /p_percent &
    scale_x_discrete(limits = c( 2000,2005, 2010,2015,2020),
                   labels = c('','2005','','2015','')) 
```
